### 2015-06-29 narrative/intro.tex
Polishing the intro

Exascale scientific discovery will be severely bottlenecked without
sufficient new research into managing and storing the large amounts of data
that will be produced during the simulation, and analyzed for months
afterwards.
%
In this project we will demonstrate a cooperative approach for storing data
where the user and the storage system work together to achieve the best data
quality and performance based on data features, storage tier characteristics,
and system state. We will investigate novel techniques to facilitate efficient
mapping of data objects, even partitioning individual variables, from the user
space onto multiple storage tiers and enable application-guided data
reductions/transformations to address capacity and bandwidth bottlenecks while
constraining the error to be within user provided bounds.
%
We will address the associated I/O and storage challenges in the context of
current and emerging storage landscapes and expedite insights into critical
scientific processes demonstrating the validity of our approach in key DOE
domains. We will research novel techniques into a Scalable
Storage Software Infrastructure integrating services from the middle-ware
layer adapted to each application with Storage System features to support the
novel data distribution both horizontally and vertically across the storage
system. The negotation between these layers will be a fundamental service 
to ensure that users will be able to save `'the best'' amount of data using the
right storage technology available at request time across the entire storage
stack.

The metric we are most interested in optimizing is time to knowledge.
Current approaches to addressing the I/O bottleneck fall into two broad
categories. Parallel file system approaches that optimize the throughput for
an entire system, and I/O middleware approaches that optimize the
performance of a single application. Both approaches have seen success but
are unlikely to overcome the major obstacles in reaching exascale. Instead
of trying to optimize throughput, we will seek to reduce the time to
knowledge. This is the most significant metric for scientific applications
where the desired outcome is not storing data, but rather executing a
knowledge extraction process on the data. Moreover, we aim to perform the
optimization not for a single application, but rather for workload of a
multi-user multi-application system environment.

Our approach leverages the expected characteristics of exascale storage
hardware. The storage layer will be partitioned into multiple heterogenous
tiers with vastly different performance characteristics. This difference
between layers will be further exacerbated by the constraints on capacity
and data lifetime within a layer. Tape archival storage will still maintain
data long term, but access to this data will be orders of magnitude slower
than the next layer and naively accessing data from archival storage will
greatly impact productivity.

To address these challenges we will We are proposing to combine the
applications knowledge from ORNL, from the ADIOS project, with the SNL
Lightweight File Systems project and storage system knowledge, the metadata
management and storage systems knowledge of UCSC, and the middleware expertise
of Rutgers to investigate how to make a data, workflow, and application aware
storage system.

@@

In this project we extend ADIOS~\cite{adios} with new APIs and novel techniques
to talk to the storage back-end. The key idea is offering the ability to
negotiate with the storage system to agree on the tradeoff between data quality
and response time.  We will test our ideas on the next generation OLCF,
SNL/LANL, and  NERSC systems.
%
In this project we will make use of the many inovations we discovered in 
the Lightweight File Systems~\cite{lwfs} (LWFS) project, where Sandia sought to
demonstrate stripping down a file system to the bare minimum features required
for any storage system. Then, additional functionality and semantics, such as
POSIX-style directories and consistency requirements, are added in using
auxiliary services separate from the stack or by layering providing a new
interface to the storage system. The project's first phase demonstrated that
using a core object storage system with integrated security providing
authentication and authorization services, customized storage interfaces could
be developed with exactly the overhead required for different applications. The
second phase of LWFS, known as Sirocco~\cite{sirocco}, is investigating how to
manage storage devices for efficient and scalable storage performance.  Sirocco
has been in process for three years and has a usable storage infrastructure
system.

>>>>

Avoiding impending bottlenecks in exascale scientific discovery will require
new research into managing and storing the large amounts of data
that will be produced by simulations and analyzed for months
afterwards.
%
In this project we will demonstrate a cooperative approach for storing data
where the user and the storage system work together to achieve the best possible data
quality and performance given data features, storage tier characteristics,
and current system state. We will investigate novel techniques to facilitate efficient
mapping of data objects, even partitioning individual variables, from the user
space onto multiple storage tiers and enable application-guided data
reductions/transformations to address capacity and bandwidth bottlenecks while
constraining the error to be within user provided bounds.
%
We will address the associated I/O and storage challenges in the context of
current and emerging storage landscapes and expedite insights into critical
scientific processes, while demonstrating the validity of our approach in key DOE
domains. We will research techniques for creating a Scalable
Storage Software Infrastructure, integrating services from the middleware
layer with Storage System features to support the
novel strategy of distributing data both horizontally and vertically across the storage
system. 
The negotation between these layers will be provided as a fundamental service, 
allowing users to easily save ``the best'' amount of data, automatically placed
throughout the entire storage stack.

The metric we are most interested in optimizing is time to knowledge.
Current approaches to addressing the I/O bottleneck fall into two broad
categories: parallel file system approaches that optimize the throughput for
an entire system, and I/O middleware approaches that optimize the
performance of a single application. Both approaches have been successful to date, but
are unlikely to overcome the remaining obstacles in reaching exascale. Instead
of trying to optimize throughput, we will seek to reduce the time to
knowledge. This is the most significant metric for scientific applications
where the desired outcome is not storing data, but rather executing a
knowledge extraction process on the data. Moreover, we aim to perform the
optimization not for a single application, but rather for workloads within a
multi-user multi-application system environment.

Our approach targets the expected characteristics of exascale storage
hardware. The storage layer will be partitioned into multiple heterogenous
tiers with vastly different performance characteristics. The layers
will be further differentiated by the constraints on capacity
and data lifetime within each layer. Tape archival storage will still maintain
data long term, but access to this data will be orders of magnitude slower
than the next layer and naively accessing data from archival storage will
greatly impact productivity.

To address these challenges we will We are proposing to combine knowledge
from a wide range of sources including ORNL applications, the ADIOS project,
the SNL Lightweight File Systems project and storage system knowledge, the metadata
management and storage systems knowledge of UCSC, and the middleware expertise
of Rutgers. We will integrate this diverse expertise to investigate how to make
a data, workflow, and application aware storage system.

@@

In this project we will extend ADIOS~\cite{adios} with new APIs and novel techniques
to coordinate with the storage back-end. The key idea is offering the ability to
negotiate with the storage system to agree on the tradeoff between data quality
and response time. We will test our ideas on the next generation OLCF,
SNL/LANL, and NERSC systems.
%
In this project we will make use of the many inovations we discovered in 
the Lightweight File Systems~\cite{lwfs} (LWFS) project, where Sandia sought to
demonstrate stripping down a file system to the bare minimum features required
for any storage system. Then, additional functionality and semantics, such as
POSIX-style directories and consistency requirements, are added in using
auxiliary services separate from the stack or by layering providing a new
interface to the storage system. The project's first phase demonstrated that
using a core object storage system with integrated security providing
authentication and authorization services, customized storage interfaces could
be developed with exactly the overhead required for different applications. The
second phase of LWFS, known as Sirocco~\cite{sirocco}, is investigating how to
manage storage devices for efficient and scalable storage performance.  Sirocco
has been in process for three years and has a usable storage infrastructure
system.
